Pre-trained encoders for dialogue: The case of laughter

### 2019-12-09 reports/ACL2020 Paper/acl2020-paper.tex
integrate references in background

\begin{document}
\maketitle
\begin{abstract}
  In this paper, we investigate the usefulness of BERT for dialogue act recognition.
  We are interested in both how useful the pre-training procedure is and 
  how well the model adapts to the dialogue domain.
  To confirm that the fine-tuned model learns to represent dialogical features, 
  we look at how it uses laughter, a phenomenon specific to dialogue.
\end{abstract}

@@

We investigate whether BERT \citep{devlinBERTPretrainingDeep2018}, a well-know pre-trained language model, is useful for dialogue.
We use dialogue act recognition  as a proxy task, since both the internal content of contributions, and their sequential structure have bearing on the task.
We also show that laughs can be helpful to disambiguate a dialogue act.

@@

\paragraph{Dialogue Act Recognition (DAR)}
The concept of a dialogue act is based on that of speech acts \citep{austinHowThingsWords2009}.
Breaking with classical semantic theory, speech act theory considers not only the propositional content of an utterance but also the actions, such as \emph{promising} or \emph{apologizing}, it carries out.
Dialogue acts extend the concept of the speech act, with a focus on the interactional nature of most speech.
DAMSL \citep{coreCodingDialogsDAMSL1997}, for example, is an influential dialogue act tagging scheme where dialogue acts are defined in part by whether they have a \emph{forward-looking} function (are expecting a response) or \emph{backward-looking} function (are in response to a previous utterance).

DAR is the the task of labeling utterances with the dialogue act they perform 
from a given set of dialogue act tags.
As with other sequence labeling tasks in NLP, some notion of context is helpful in DAR.
One of the first performant machine learning models for DAR was a Hidden Markov Model that used various lexical and prosodic features as input \citep{stolckeDialogueActModeling2000}.
Most successful neural approaches also model some notion of context, 
for example with a recurrent model \citep[e.g.,][]{botheContextbasedApproachDialogue2018} or conditional random field \citep[e.g.,][]{chenDialogueActRecognition2017}.

\begin{itemize}
  % related because of the DAR sequence model
  \item \citet{kalchbrennerRecurrentConvolutionalNeural2013} -- DAR model similar to our architecture. Uses an RNN ``discourse model'' and a CNN similar to our baseline encoder
  \item \citet{botheContextbasedApproachDialogue2018} -- uses an RNN for context (recently SOTA on SWDA)
  \item \citet{tranHierarchicalNeuralModel2017} -- RNNs at both the utteranec and dialogue level. Also attention. Very similar to our architecture.
  \item \citet{tranPreservingDistributionalInformation2017} -- Sequential DA prediction with uncertainty propegation
  % cf.
  \item \citet{ortegaNeuralbasedContextRepresentation2017} -- SWDA DAR with attention-based context representations and CNN utterance representations
  % related because of the utt encoder model
  \item \citet{khanpourDialogueActClassification2016} -- Deep LSTM utt representation. No notion of context. Finds word vectors are useful.
  \item \citet{shenNeuralAttentionModels2016} -- Uses attention for utt encoding in DAR. No notion of context, I think.
  % related because of methodology
  \item \citet{pragstVectorRepresentationUtterances2018} -- dialogue vector models. Uses vector-manual DA tag correspondance to argue for semantic relevance of their vector models (related to how we use DA tagging to verify BERT usefulness)
  % ???
  \item \citet{chenDialogueActRecognition2017} -- SOTA? SWDA DAR using attention and CRF over utterances
\end{itemize}

\paragraph{Transfer learning for NLP}
Transfer learning is are techniques that allow a model trained on one task---often unsupervised---to be applied to another. 
Since annotating natural language data is expensive, there is a lot of interest in transfer learning for natural language processing. 
The use of pre-trained word vectors \citep[e.g.,][]{mikolovDistributedRepresentationsWords2013;penningtonGloveGlobalVectors2014} is a ubiquitous example of transfer learning in NLP.
Recently, large neural models trained on various unsupervised tasks have found success in producing sentence-level and contextual word representations.

BERT \citep{devlinBERTPretrainingDeep2018} is one such model. 
It is a large multi-layer transformer, pre-trained on two unsupervised tasks: \emph{masked token prediction} and \emph{next sentence prediction}.
In masked token prediction, some percentage of words are randomly replaced with a \texttt{[MASK]} token, and the model is trained to predict the identity of these tokes based on the context sentence.
In next sentence prediction, the model is give two sentences and trained to predict whether the second sentence follows the first in the original text, or if it was randomly chosen from elsewhere in the corpus.
After pre-training, BERT can be applied to a supervised task by adding additional un-trained layers that take the hidden state of one or more of BERT's layers as input.

Transfer learning schemes for pre-trained language models such as BERT draw on one or more of the following three adaptation strategies \citet{sunHowFineTuneBERT2019}:
\emph{further pre-training}, in which the model is trained in an un-supervised way, similar to its initial training scheme, but on data that is in-domain for the target task; 
\emph{single-task fine-tuning}, in which the model's encoder layers are optimized during training for the target task;
and \emph{multi-task fine-tuning}, in which models with shared encoder layers are jointly trained on several related tasks.
In this paper, we investigate the first two methods and their combination, as applied to DAR.

@@

\begin{itemize}
  \item \citet{cerisaraEffectsUsingWord2vec2017} -- standard pre-trained word embeddings do not help for DAR in any of three tested languages (English, French, Czech)
    % BERT for dialogue (not nesc. DAR)
  \item \citet{chenSemanticallyConditionedDialog2019a} -- uses BERT for response generation
  \item \citet{mehriPretrainingMethodsDialog2019} -- looks at various dialogue tasks using pre-training, including DAR and using BERT
  \item \citet{baoPLATOPretrainedDialogue2019} -- pre-training for dialogue generation
  \item \citet{vigComparisonTransferLearningApproaches} -- Transfer learning for response selection
\end{itemize}

\paragraph{Contributions}
The main results of this paper have to do with to the usefulness of language model fine-tuning for DAR, the impact of laughter on DAR, and the relationship between the two.
\begin{itemize}
  \item Laughter is useful for dialogue act recognition, and its impact varies across different dialogue acts (\S\ref{sec:experiment1}).
  \item During fine-tuning, BERT learns to represent laughter, a dialogical feature not seen in pre-training (\S\ref{sec:experiment2} and \S\ref{sec:experiment3}).
  \item Standard BERT pre-training is useful for DAR, but the model performs poorly without fine-tuning (\S\ref{sec:experiment2})
  \item Further pre-training with in-domain data shows promise for dialogue (\S\ref{sec:experiment3}), but further investigation with larger dialogue corpora is required.
\end{itemize}

\section{Data}
We perform experiments on the Switchboard Dialogue Act Corpus (SWDA), which is a subset of the larger Switchboard corpus, and the dialogue act-tagged portion of the AMI Meeting Corpus (AMI-DA). 
SWDA is tagged with a set of 220 dialogue act tags which, following  \citet{jurafskySwitchboardSWBDDAMSLShallowDiscourseFunction1997a} we cluster into a smaller set of 42 tags.
AMI uses a smaller tagset of 16 dialogue acts \citep{GuidelinesDialogueAct2005}.

@@

Confusion matrices (figure \ref{fig:swda-cm}) provide some food for thought. Most of the misclassifications fall into the majority classes, such as \emph{sd} (Statement-non-opinion), in the left edge of the matrix. However there are some important exceptions, such as \emph{rhetorical questions}, that are misclassified as other forms of questions due to their surface question-like form.
Importantly, laughter helpes to classify rhetorical questions correctly, this is because in a conversation it can be used as a device to cancel seriousness \citep{ginzburg2015understanding}, in this case, of a question.
Therefore questions, like the one we show in the example in table~\ref{table:example-qh}, are easier to disambiguate with the help of laughter.
\begin{table}
      \small
  \centering
  \begin{tabularx}{\linewidth}{llL}
    \toprule
    Speaker & DA & Utterance \\ \midrule
        \multicolumn{3}{c}{\emph{(Talking about hobbies)}}\\
    B   & sd	&  Um, as far as spare time, they talked about, \\
    B	& \% & I don't, + I think, \\
    B	& qh & who has any spare time \texttt{<laughter>}? \\
    A	& x & \texttt{<laughter>}.\\
             \bottomrule
  \end{tabularx}
  \caption{Example from the SWDA corpus (sw3735). B's contribution \emph{qh} (Rhetorical question) is misinterpreted as \emph{qw} (Wh-question) by the BERT model without laughs in training data. }
  \label{table:example-qh}
\end{table}

  \begin{figure}
  \centering
  \includegraphics[width=1.0\linewidth]{img/swda-cm.pdf}
  \caption{Confusion matrices for \texttt{BERT-NL} (top) vs \texttt{BERT-L} (bottom); SWDA corpus. Solid line shows the improvement in classification of rhetorical questions.}
    \label{fig:swda-cm}
  \end{figure}

>>>>

\begin{document}
\maketitle
\begin{abstract}
  The degree to which language models pre-trained on text data are useful for dialogue is an open question.
  In this paper, we investigate the usefulness of BERT for dialogue act recognition.
  We are interested in both how useful the pre-training procedure is and 
  how well the model adapts to the dialogue domain.
  To confirm that the fine-tuned model learns to represent dialogical features, 
  we look at how it uses laughter, a phenomenon specific to dialogue.
\end{abstract}

@@

In this work, we investigate whether BERT \citep{devlinBERTPretrainingDeep2018}, a well-know pre-trained language model, is useful for dialogue.
We use dialogue act recognition  as a proxy task, since both the internal content of contributions, and their sequential structure have bearing on the task.
In this work we also show that laughs can be helpful to disambiguate a dialogue act.

@@

\subsection{Dialogue Act Recognition}
The concept of a dialogue act is based on that of speech acts \citep{austinHowThingsWords2009}.
Breaking with classical semantic theory, speech act theory considers not only the propositional content of an utterance but also the actions, such as \emph{promising} or \emph{apologizing}, it carries out.
Dialogue acts extend the concept of the speech act, with a focus on the interactional nature of most speech.
DAMSL \citep{coreCodingDialogsDAMSL1997}, for example, is an influential dialogue act tagging scheme where dialogue acts are defined in part by whether they have a \emph{forward-looking} function (are expecting a response) or \emph{backward-looking} function (are in response to a previous utterance).

DAR is the the task of labeling utterances with the dialogue act they perform 
from a given set of dialogue act tags.
As with other sequence labeling tasks in NLP, some notion of context is helpful in DAR.
One of the first performant machine learning models for DAR was a Hidden Markov Model that used various lexical and prosodic features as input \citep{stolckeDialogueActModeling2000}.

Most successful neural approaches also model some notion of context.
We follow previous work that models dialogue act sequences with an RNN over utterances \citep{kalchbrennerRecurrentConvolutionalNeural2013,tranHierarchicalNeuralModel2017,botheContextbasedApproachDialogue2018}.
Our model predicts the most likely tag given the hidden state of the RNN at each position, though other work has shown that DAR benefits from more sophistacted approaches, such as conditional random field \citep{chenDialogueActRecognition2017} or uncertainty propegation \citep{tranPreservingDistributionalInformation2017}.

%\begin{itemize}
  %\item \citet{kalchbrennerRecurrentConvolutionalNeural2013} -- DAR model similar to our architecture. Uses an RNN ``discourse model'' and a CNN similar to our baseline encoder
  %\item \citet{botheContextbasedApproachDialogue2018} -- uses an RNN for context (recently SOTA on SWDA)
  %\item \citet{tranHierarchicalNeuralModel2017} -- RNNs at both the utteranec and dialogue level. Also attention. Very similar to our architecture.
  %\item \citet{tranPreservingDistributionalInformation2017} -- Sequential DA prediction with uncertainty propegation
  %\item \citet{ortegaNeuralbasedContextRepresentation2017} -- SWDA DAR with attention-based context representations and CNN utterance representations
  %\item \citet{shenNeuralAttentionModels2016} -- Uses attention for utt encoding in DAR. No notion of context, I think.
  % related because of methodology
  %\item \citet{pragstVectorRepresentationUtterances2018} -- dialogue vector models. Uses vector-manual DA tag correspondance to argue for semantic relevance of their vector models (related to how we use DA tagging to verify BERT usefulness)
  %\item \citet{chenDialogueActRecognition2017} -- SOTA? SWDA DAR using attention and CRF over utterances
%\end{itemize}

\paragraph{Transfer learning for NLP}
Transfer learning techniques allow a model trained on one task---often unsupervised---to be applied to another. 
Since annotating natural language data is expensive, there is a lot of interest in transfer learning for natural language processing. 
Word vectors \citep[e.g.,][]{mikolovDistributedRepresentationsWords2013,penningtonGloveGlobalVectors2014} are a ubiquitous example of transfer learning in NLP.
We note, however, that pre-trained word vectors are not always useful when applied to dialogue domains \citet{cerisaraEffectsUsingWord2vec2017}.

Recently, neural models trained on large datasets with various unsupervised tasks have found success creating sentence-level and contextual word representations. 
BERT, a multi-layer tranformer, is one such model \citep{devlinBERTPretrainingDeep2018} . 
BERT is pre-trained on two unsupervised tasks: \emph{masked token prediction} and \emph{next sentence prediction}.
In masked token prediction, some percentage of words are randomly replaced with a \texttt{[MASK]} token, and the model is trained to predict the identity of these tokes based on the context sentence.
In next sentence prediction, the model is give two sentences and trained to predict whether the second sentence follows the first in the original text, or if it was randomly chosen from elsewhere in the corpus.
After pre-training, BERT can be applied to a supervised task by adding additional un-trained layers that take the hidden state of one or more of BERT's layers as input.

There is some preivous work applying BERT to dialogue.
\citet{baoPLATOPretrainedDialogue2019,chenSemanticallyConditionedDialog2019a} and both use BERT for dialogue generation tasks.
Similarly \citet{vigComparisonTransferLearningApproaches2019} finds BERT useful for selecting a response from a list of candidate responses in a dialogue.
Finally, \citet{mehriPretrainingMethodsDialog2019} evaluates BERT in various dialogue tasks including DAR, and find that a model incorporating BERT outperforms a baseline model.

\paragraph{Fine-tuning vs.~further in-domain pre-training}
Transfer learning schemes for pre-trained language models such as BERT draw on one or more of the following three adaptation strategies \citet{sunHowFineTuneBERT2019}:
\emph{further pre-training}, in which the model is trained in an un-supervised way, similar to its initial training scheme, but on data that is in-domain for the target task; 
\emph{single-task fine-tuning}, in which the model's encoder layers are optimized during training for the target task;
and \emph{multi-task fine-tuning}, in which models with shared encoder layers are jointly trained on several related tasks.
In this paper, we investigate the first two methods and their combination, as applied to DAR.

@@

%\begin{itemize}
  %\item \citet{cerisaraEffectsUsingWord2vec2017} -- standard pre-trained word embeddings do not help for DAR in any of three tested languages (English, French, Czech)
  %\item \citet{khanpourDialogueActClassification2016} -- Deep LSTM utt representation. No notion of context. Finds word vectors are useful.
     %BERT for dialogue (not nesc. DAR)
  %\item \citet{chenSemanticallyConditionedDialog2019a} -- uses BERT for response generation
  %\item \citet{mehriPretrainingMethodsDialog2019} -- looks at various dialogue tasks using pre-training, including DAR and using BERT
  %\item \citet{baoPLATOPretrainedDialogue2019} -- pre-training for dialogue generation
  %\item \citet{vigComparisonTransferLearningApproaches} -- Transfer learning for response selection
%\end{itemize}

\section{Data}\label{sec:data}
We perform experiments on the Switchboard Dialogue Act Corpus (SWDA), which is a subset of the larger Switchboard corpus, and the dialogue act-tagged portion of the AMI Meeting Corpus (AMI-DA). 
SWDA is tagged with a set of 220 dialogue act tags which, following  \citet{jurafskySwitchboardSWBDDAMSLShallowDiscourseFunction1997a} we cluster into a smaller set of 42 tags.
AMI uses a smaller tagset of 16 dialogue acts \citep{GuidelinesDialogueAct2005}.

@@

Confusion matrices (figure \ref{fig:swda-cm}) provide some food for thought.
% TODO: explain why most of misclassifications are sd (majority class)
One of the most obvious things that can be observed is rhetorical questions: supposedly, laughter can be used as a device to cancel seriousness \citep{ginzburg2015understanding} of the question.
Therefore it can help to disambiguate such questions like the one we show in example in table~\ref{table:example-qh}.
\begin{table}
      \small
  \centering
  \begin{tabularx}{\linewidth}{llL}
    \toprule
    Speaker & DA & Utterance \\ \midrule
        \multicolumn{3}{c}{\emph{(Talking about hobbies)}}\\
    B   & sd	&  Um, as far as spare time, they talked about, \\
    B	& \% & I don't, + I think, \\
    B	& qh & who has any spare time \texttt{<laughter>}? \\
    A	& x & \texttt{<laughter>}.\\
             \bottomrule
  \end{tabularx}
  \caption{Example from the SWDA corpus (sw3735). B's contribution \emph{qh} (Rhetorical question) is misinterpreted as \emph{qw} (Wh-question) by the BERT model without laughs in training data. }
  \label{table:example-qh}
\end{table}

  \begin{figure}
  \centering
  \includegraphics[width=0.9\linewidth]{img/swda-cm.pdf}
  \caption{Confusion matrices for \texttt{BERT-NL} (top) vs \texttt{BERT-L} (bottom); SWDA corpus. Solid line shows the improvement in classification of rhetorical questions.}
    \label{fig:swda-cm}
  \end{figure}
